---
title: "Dataset Preparation"
author: "Noelle Schenk"
date: "4/29/2022"
output: html_document
---

# Functions dataset

```{r}
library(data.table)
library(reshape2)
library(BetaDivMultifun)
# phylo traits course datasets
path_to_data <- "/run/user/1000/gvfs/smb-share:server=nas-ips,share=ips/groups/planteco/PROJECTS/Exploratories Synthesis/Courses_Workshops/2022_PhyloTrait_Workshop/Data/"
# functions dataset for metadata "funtab"
path_to_functions <- "/run/user/1000/gvfs/smb-share:server=nas-ips,share=ips/groups/planteco/PROJECTS/Exploratories Synthesis/Data/Grassland_functions/27087_grassland_functions_bexis/27087_25_Dataset/"
path_to_proj <- "/run/user/1000/gvfs/smb-share:server=nas-ips,share=ips/groups/planteco/PROJECTS/Exploratories Synthesis/Research Projects/BetaDivMultifun/"
```

1. List of available grassland functions
2. Open-access dataset of grassland functions

Make list of available grassland functions
```{r, eval = F}
funtab <- fread(paste(path_to_functions, "UPDATE29-04-22_synthesis_grassland_function_metadata_ID27087.csv", sep = ""))
write.csv(unique(funtab[, .(ColumnName, ground)]), "funnames.csv")
# see the excel table made from this one
```


## 17 selected functions which can go together in an analysis
Clean functions to open-access dataset of grassland functions
```{r}
imputed_grlfuns <- readRDS(paste(path_to_proj, "/data_assembly/output_data/imputed_and_compound_grlfuns.rds", sep = ""))
imputed_grlfuns[, herbivory.20172018 := NULL] # only non-open access function is herbivory.20172018
ncol(imputed_grlfuns)
fwrite(imputed_grlfuns, "27087_17_selected_grassland_functions.csv")
```

## Biomass over years
```{r}
grlfuns <- fread(paste(path_to_functions, "bexis_to_wide_format_output.csv", sep = ""))
grlfuns <- fread(paste(path_to_functions, "27087_25_data.csv", sep = ""))
grlfuns <- grlfuns[, .(Plotn, Year, Biomass)]
grlfuns <- grlfuns[Year %in% seq(2008, 2017, 1)]
fwrite(grlfuns, "27087_grl_functions_biomass_yearly.csv", sep = ";")
```



# Covariates

## LUI
```{r}
### create delta LUI dataset ----
deltaLUI <- fread(paste(path_to_data, "deltaLUI.csv", sep = ""))
deltaLUI[, YEAR := 0000]
deltaLUI <- deltaLUI[, .(Plotn, YEAR, deltaLUI, deltaGstd, deltaMstd, deltaFstd)]
### create LUI dataset ----
LUI <- fread(paste(path_to_data, "LUIData/LUI_new components set_global_separately_2022-04-29.txt", sep = ""))
# clean LUI
LUI[, EXPLO := NULL]
LUI$YEAR <- as.numeric(sub("\\)", "", sub("separately\\(", "", LUI$YEAR, perl = F)))
# change PlotIDs
setnames(LUI, old = c("PLOTID"), new = c("Plot"))
a <- c(paste("AEG", stringr::str_pad(seq(1, 50), 2, pad = "0"), sep = ""),
  paste("HEG", stringr::str_pad(seq(1, 50), 2, pad = "0"), sep = ""),
  paste("SEG", stringr::str_pad(seq(1, 50), 2, pad = "0"), sep = ""))
b <- c(paste("AEG", seq(1, 50), sep = ""),
  paste("HEG", seq(1, 50), sep = ""),
  paste("SEG", seq(1, 50), sep = ""))
usefulplotids <- data.table(Plot = b, Plotn = a)
LUI <- merge(usefulplotids, LUI, by = "Plot")
LUI[, Plot := NULL]
# calc year 0 : mean over all years
year0 <- data.table(merge(merge(merge(
  aggregate(LUI ~ Plotn, LUI, mean),
      aggregate(G_STD ~ Plotn, LUI, mean), by = "Plotn"),
      aggregate(M_STD ~ Plotn, LUI, mean), by = "Plotn"),
      aggregate(F_STD ~ Plotn, LUI, mean), by = "Plotn"))
year0[, YEAR := 0]
LUI <- rbindlist(list(LUI, year0), use.names = T)
# merge delta LUI with LUI dataset
combLUI <- merge(LUI, deltaLUI, by = c("Plotn", "YEAR"), all = T)
setnames(combLUI, old = c("YEAR"), new = c("Year"))
fwrite(combLUI, file = "combinedLUI.csv", sep = ";")
```

## covariates
plot isolation, longitude and latitude of plots
soil covariates : run parts of the BetaDivMultifun scripts : "clean_and_load_soil_covariates.Rmd", "calc_covariates.Rmd"
```{r}
covariates <- fread(paste(path_to_data, "raw_soil_and_landscape_covariates.csv", sep =""))
```

raw soil variables + 1st axis of a PCA
Calculate PCA of soil variables
```{r}
soilpca_output <- pca_from_glsoil$x[, 1:3]
soilpca_output <- data.table(cbind("Plotn" = rownames(soilpca_output), soilpca_output))
colnames(soilpca_output)[-1] <- paste("soil", colnames(soilpca_output)[-1], sep = "")
# merge to covariates
covariates <- merge(covariates, soilpca_output, by = "Plotn", all = T)
fwrite(covariates, "soil_and_landscape_covariates.csv", sep = ";")
```

